---
title: "Data Wrangling using dplyr"
author: "Florian Pätzold, 977687"
output: rmarkdown::github_document
---

```{r echo=F}
knitr::opts_chunk$set(
  warning = FALSE, # supress warnings per default 
  message = FALSE  # supress messages per default 
)
```

```{r preprocessing, echo=F}

## Preprocessing

# let's load in the libraries we need for the following session. 
library(tidyverse)

#devtools::install_github("michael-franke/aida-package")
library(aida)

# and our dataset, let's call it dolphin
dolphin <- aida::aidata

```

## Exercises
### Exercise 1

(a) Take the dolphin data set and store a reduced variant of it as `dolphin_reduced`. The new data frame should contain only the following columns: `RT`, `AUC`, `group`, and `exemplar`

```{r x1a}
dolphin_reduced <- 
  dolphin %>% 
  dplyr::select(RT, AUC, group, exemplar)

head(dolphin_reduced)
```

(b) We are for now only interested in those data that have whales as the exemplar. `filter()` only those rows and store them in a new dataframe called `whales_only`.

```{r x1b}
whales_only <- 
  dolphin_reduced %>% 
  filter(exemplar == "whale")

head(whales_only)
```

(c) Now filter for only those data that have RTs below 1500ms.

```{r x1c}
whales_maxRT <- 
  whales_only %>% 
  filter(RT < 1500)

head(whales_maxRT)
```

(d) We don't like that `AUC` is unstandardized. Use `mutate()` to create a new vector that represents scaled AUC values (scaling is achieved by the function `scale()`).

```{r x1d}
whales_standardized <- whales_maxRT %>% 
  mutate(AUC_scaled = as.numeric(scale(AUC)))

head(whales_standardized)
```

(e) Calculate the mean scaled AUC ratings for both both groups.

```{r x1e}
whales_standardized <- 
  whales_standardized %>% 
  group_by(group) %>% 
  summarize(mean_AUC = mean(AUC_scaled))

head(whales_standardized)
```

(f) Do all of the above (a-e) in one pipeline.

```{r x1f}
AUC_per_group <-
  dolphin %>%
  dplyr::select(RT, AUC, group, exemplar) %>%
  filter(exemplar == "whale", RT < 1500) %>% 
  mutate(AUC_scaled = as.numeric(scale(AUC))) %>% 
  group_by(group) %>% 
  summarize(mean_AUC = mean(AUC_scaled))

AUC_per_group
```


## HOMEWORK
### please upload your solutions to studIP as an .Rmd file

### Exercise 2 

(a) Take the dolphin data set and store a reduced variant of it. The new data frame should contain only the columns `condition`, `group`, and `xpos_flips`, `correct`. And within the `correct` vector, we are only interested in the correct trials (= 1). Filter accordingly.

```{r x2a}
dolphin_reduced <- 
  dolphin %>%
  dplyr::select(condition, group, xpos_flips, correct) %>% 
  filter(correct == 1)

head(dolphin_reduced)
```

(b) Create an aggregated data frame that contains the mean `xpos_flips` value and the standard deviation for `group` and `condition`.

```{r x2b}
dolphin_aggregated <-
  dolphin_reduced %>% 
  group_by(group, condition) %>% 
  summarize(mean_xpos_flips = mean(xpos_flips),
            sd_xpos_flips = sd(xpos_flips))

head(dolphin_aggregated)
```

(c) Rename the new vectors for the mean xflips and their sd to `xflips_mean` and `xflips_sd`

```{r x2c}
dolphin_aggregated <- 
  dolphin_aggregated %>% 
  rename(xflips_mean = mean_xpos_flips,
         xflips_sd = sd_xpos_flips)

head(dolphin_aggregated)
```

(d) Do all of the above (a-c) in one pipeline.

```{r 2xd}
dolphin_cleaned <-
  dolphin %>% 
  dplyr::select(condition, group, xpos_flips, correct) %>% 
  filter(correct == 1) %>% 
  group_by(group, condition) %>%
  # Note, that I rename the columns already in the 
  # summarize function in order to reduce redundancy.
  summarize(xflips_mean = mean(xpos_flips),
            xflips_sd = sd(xpos_flips))

head(dolphin_cleaned)
```




