---
title: "Data Visualization using ggplot2"
author: "Florian Pätzold, 977687"
output: html_document
---

```{r echo=F}
knitr::opts_chunk$set(
  warning = FALSE, # supress warnings per default 
  message = FALSE  # supress messages per default 
)
```

```{r preprocessing, echo=F}

## Preprocessing

# let's load in the libraries we need for the following session. 
library(tidyverse)
library(ggbeeswarm)

#devtools::install_github("michael-franke/aida-package")
library(aida)

# our dataset
dolphin <- aida::aidata

```

## Exercises

### Exercise 1

Take the scatter plot below as a departure point. It plots AUC (area-under-the-curve) against MAD (maximal absolute deviation). 

```{r x1, warning = F, fig.width = 8, fig.align="center"}

ggplot(dolphin, aes(x = MAD, y = AUC)) +
  geom_point() +
  ggtitle("a scatter plot") +
  theme_aida()

```

a) Change both the x-axis and the y-axis title to sensible and informative titles.

```{r x1a, fig.width = 8, fig.align="center"}

ggplot(dolphin, aes(x = MAD, y = AUC)) +
  geom_point() +
  ggtitle("a scatter plot") +
  xlab("\n Size of Competition effect (MAD)") +
  ylab("Size of typicality effect (AUC) \n") +
  theme_aida()

```

b) Change the plot title to something informative.

```{r x1b, fig.width = 8, fig.align="center"}

ggplot(dolphin, aes(x = MAD, y = AUC)) +
  geom_point() +
  ggtitle("Area under the curve plotted onto the maximum absolute deviation") +
  xlab("\n Size of Competition effect (MAD)") +
  ylab("Size of typicality effect (AUC) \n") +
  theme_aida()

```

c) Change the scaling of the x-axis to display only MAD values between -500 and 500

```{r x1c, fig.width = 8, fig.align="center"}

ggplot(dolphin, aes(x = MAD, y = AUC)) +
  geom_point() +
  ggtitle("Area under the curve plotted onto the maximum absolute deviation") +
  xlab("\n Size of Competition effect (MAD)") +
  ylab("Size of typicality effect (AUC) \n") +
  scale_x_continuous(limits = c(-500,500)) +
  theme_aida()

```

d) Add a color aesthetic for the predictor `group`.

```{r x1d, fig.width = 8, fig.align="center"}

ggplot(dolphin, aes(x = MAD, y = AUC, color = group)) +
  geom_point() +
  ggtitle("Area under the curve plotted onto the maximum absolute deviation") +
  xlab("\n Size of Competition effect (MAD)") +
  ylab("Size of typicality effect (AUC) \n") +
  scale_x_continuous(limits = c(-500,500)) +
  theme_aida()

```

### HOMEWORK ###
## please upload your solutions to studIP as an .Rmd file

## Exercise 2

a) Plot AUC values as a function of group in a density plot (`geom_density`).

```{r x2a, fig.width = 8, fig.align = "center"}

ggplot(dolphin, aes(x = AUC, color = group, fill = group)) +
  geom_density()

```

b) Color code the curves (the border and the filling) as follows: "orange" for the click group and "black" for the touch group.

```{r x2b, fig.width = 8, fig.align = "center"}

ggplot(dolphin, aes(x = AUC, color = group, fill = group)) +
  geom_density() +
  scale_colour_manual(values = c("orange", "black"), aesthetics = c("colour", "fill"))

```

c) Make the density curves semi-transparent with the `alpha` argument.

```{r x2c, fig.width = 8, fig.align = "center"}

ggplot(dolphin, aes(x = AUC, color = group, fill = group)) +
  geom_density(alpha = 0.5) +
  scale_colour_manual(values = c("orange", "black"), aesthetics = c("colour", "fill"))

```

d) Add the aida_theme to the plot.

```{r x2d, fig.width = 8, fig.align = "center"}

ggplot(dolphin, aes(x = AUC, color = group, fill = group)) +
  geom_density(alpha = 0.5) +
  scale_colour_manual(values = c("orange", "black"), aesthetics = c("colour", "fill")) +
  theme_aida()

```

e) BONUS: Add the mean values for both groups into the density plot as a line.

```{r x2e, fig.width = 8, fig.align = "center"}

# Wrangle data frame to necessary values.
AUC_means <- 
  dolphin %>%
  dplyr::select(group, AUC) %>% 
  group_by(group) %>% 
  summarise(mean_AUC = mean(AUC, na.rm=T))

AUC_means

# Extract the numbers in question for a better overview 
# by accessing the data frame column "mean_AUC".
click_mean <- AUC_means$mean_AUC[1]
touch_mean <- AUC_means$mean_AUC[2]
  
ggplot(dolphin, aes(x = AUC, color = group, fill = group)) +
  geom_density(alpha = 0.5) +
  scale_colour_manual(values = c("orange", "black"), aesthetics = c("colour", "fill")) +
  # Let's add the mean values as vertical lines.
  geom_vline(xintercept = click_mean, color = "orange") +
  geom_vline(xintercept = touch_mean, color = "black") +
  # Let's also add some label descriptions.
  labs(title = "A density plot of AUC values.", 
       x = "\n Area under the curve", 
       y = "Density \n", 
       caption = "\n The vertical lines indicate the mean 
                  value of the AUC for each group.") +
  theme_aida()

```


